Icicle: fuse your queries

### 2016-04-20 2016fhpc/section/XX-Related.tex
notes about GVN

``most compilers restrict themselves to common subexpression elimination within basic blocks or extended basic blocks''
\cite{debray1992compiler}

@@

Multi-query data analysis\cite{andrade2003efficient} describes a similar problem, where they have multiple queries over the same data that run periodically.
By grouping the multiple queries together, more optimisations can be performed than when treating the queries separately.
This is a lot closer, and performs CSE on expressions like @x = sum(y)@, where @sum@ is a built-in aggregate function.
However, this is only for a small set of built-in aggregates, and does not deal with mutually recursive

>>>>

The most closely related transforms are induction variable elimination\cite{shivers1988control} and global value numbering\cite{rosen1988global}.
Induction variable elimination finds and removes accumulators that are linear functions of the iteration number.
This does not deal with non-linear functions such as sum, or mutually recursive accumulators.

Global value numbering works on an SSA form and is much more general than induction variable elimination.
Early versions such as Rosen et al\cite{rosen1988global} only worked on extended basic blocks with backedges removed.
Alpern et al\cite{alpern1988detecting} introduced cases for specific loop backedges, where a fixpoint is performed to find the congruence sets of loops.
More recently, Gulwani and Necula\cite{gulwani2004polynomial} improved upon this, supporting removal of more expressions, and imposing a polynomial time bound.
Global value numbering for loops can perform all optimisations here.

Common subexpression elimination is very closely related, but most imperative compilers only perform common subexpression elimination on straight-line computations with no control flow\cite{debray1992compiler}, or on loop invariant expressions\cite{bodik1998complete}.
Common subexpression elimination for functional languages would be able remove lone expressions, but does not deal with mutually recursive folds.

@@

Multi-query data analysis\cite{andrade2003efficient} describes a similar problem, where they have multiple queries over the same data that run periodically.
By grouping the multiple queries together, more optimisations can be performed than when treating the queries separately.
This is a lot closer, and performs CSE on expressions like @x = sum(y)@, where @sum@ is a built-in aggregate function.
However, this is only for a small set of built-in aggregates, and does not deal with mutually recursive definitions.

The polyhedral model\cite{benabderrahmane2010polyhedral} analyses data dependencies in loops.
For a given iteration, the polyhedral model computes the iterations which must be executed before the given iteration.
By knowing the dependent iterations, one can restructure the loop so that the iterations are executed in a different order, while still performing any dependencies in order.
For example, a loop iterating over @i@, with loop body @A[i] = A[i-10]@ depends on the tenth previous iteration.
This loop could be executed in an order like @0, 1, 2,...@, but it could also stride by ten: @0, 10, 20..., 1, 11, 21,...@.
This allows fusion of any two loops if the intersection of the two loops' iteration spaces is not empty.
In our case the iteration order is fixed, as the program cannot control the order in which stream elements are received.
The polyhedral model may still be applicable to streaming computations where one could introduce certain-sized buffers, then process the buffer in parallel.
In this case the polyhedral model may expose more opportunities for duplicate elimination, but does not itself remove duplicates.
